---
title: Report Tutorial
date: today
author: 'Australian Institute of Marine Science'

license: "CC BY"
crossref:
  lof-title: "List of Figures"

filters:
  - authors-block

format:
  html:
    toc: true
#    bibliography: references.bib
    fig-width: 8
    fig-height: 4
    code-fold: true
    embed-resources: true
    
#Tonga 93391 2017-2024
#Palau 1705 2002-2016
#PNG 65823 2020-2025
#Vanuatu 1707 2014-2018
#Fiji 1725 2006-2021
#Samoa 43039 2022-2023
---

::: lead
In this vignette, we will explore how to generate a report using the functions from the reefcloudReporting package and the data from the ReefCloud Public Dashboard from the Palau region.
:::

```{r}
#| label: loadPackages
#| include: false
library(tidyverse)
library(sf)
library(knitr)
library(patchwork)
#library(reefcloudReporting)
```

# Accessing ReefCloud Data

ReefCloud data on the [Public Dashboard](https://reefcloud.ai/dashboard/ "ReefCloud Public Dashboard") comes from the projects that users have uploaded on ReefCloud itself. These are shared as a summary of the information which are collated into sites, countries and regions, all of which are searchable and identified using a **Tier ID** that is used for most of the functions.

Here, we use the **Tier ID: 1705** to call for information from Palau to generate a table of key regional summary statistics about the region (@tbl-readRegionalSummary).

```{r}
#| label: tbl-readRegionalSummary
#| tbl-cap: "Summary of key tier characteristics"
#| warning: false
source("./R/get_regional_summary.R")

info <- get_regional_summary(1705) 

info_tbl <- tibble(
  field = names(info),
  value = sapply(info, function(x) {
    if (length(x) > 1) paste(x, collapse = ", ") else x
  })
)

knitr::kable(info_tbl) 
```

We then extract the site information from the same Tier ID, which uses the get_regional_summary function to extract Site IDs and then summarises the site level information. This can take a while if there are many sites within the Tier ID.

This gives us a summary of the sites found within the Tier ID (@tbl-readSiteSummary).

```{r}
#| label: tbl-readSiteSummary
#| tbl-cap: "Summary of key survey sites characteristics"
#| warning: false
source("./R/get_site_summary.R")

sites <- get_site_summary(1705)

sites_tbl <- knitr::kable(sites[1:10,])

sites_tbl
```

With the sites and their IDs from the get_site_summary function, we can then extract the different benthic covers for each of these sites. We read the site IDs into get_benthic_cover function to get benthic groupings and cover that we can use for other plots. 

Here, because of the number of rows of data, we just show the first 10 rows of the benthic cover table (@tbl-getCover).

```{r}
#| label: tbl-getCover
#| tbl-cap: "Mean benthic cover of survey sites within the Region"
#| warning: false
source("./R/get_benthic_cover.R")

cover <- get_benthic_cover(sites$site_id)

cover_tbl <- knitr::kable(cover[1:10,])

cover_tbl
```

# Generating Report Figures

The data we have accessed through the ReefCloud API have been converted into usable formats for visualisations. We can then proceed to plot some relevant figures that may be useful to better understand the patterns we are seeing.

First, we generate a map of the study region using generate_map. 

```{r}
#| label: fig-plotMap
#| fig-cap: "Map of region and surrounding reef sites"
#| warning: false
source("./R/generate_map.R")

# Need to redo the codes for this because its taking up too much memory with the rest of the doc.

#map_plot <- generate_map(1705)

#map_plot
```

We can also plot the Hard Coral cover for each site at the latest year. This gives us an idea of how the hard coral cover varies across the different site and can highlight sites that may be doing poorly.

For this plot, there are a few options to specify other than tier ID which we have been using previously. This includes year, cover_type and depth. Do note the options in specifying these variables. We then call the plot_site_cover function to make the plot for Hard Coral cover (@fig-plotHCCover). 

```{r}
#| label: fig-plotHCCover
#| fig-cap: "Site-level Hard Coral cover across the region"
#| warning: false
source("./R/plot_site_cover.R")

HC_cover_plot <- plot_site_cover(1705, year = 2016, cover_type = "HARD CORAL", depth = "shallow")

HC_cover_plot$plot

HC_cover_plot$df.sum
```

As an example. we also want to use plot_site_cover to plot the macroalgae cover (@fig-plotMACover).

```{r}
#| label: fig-plotMACover
#| fig-cap: "Site-level Macroalgae cover across the region"
#| warning: false
source("./R/plot_site_cover.R")

MA_cover_plot <- plot_site_cover(1705, year = 2016, cover_type = "MACROALGAE", depth = "shallow")

MA_cover_plot$plot

MA_cover_plot$df.sum
```

We can also present the information in two donut plot using plot_donut to visualise all the sites as a whole.

```{r}
#| label: fig-plotDonut
#| fig-cap: "Proportion of sites within different cover bands in the region"
#| warning: false
source("./R/plot_donut.R")

HC_donut_plot <- plot_donut(1705, cover_type = "HARD CORAL")

MA_donut_plot <- plot_donut(1705, cover_type = "MACROALGAE")

HC_donut_plot + MA_donut_plot
```

# Discussing the Results

With each of the figures, we can then write some text and description to help substantiate the points we are making. Figures on their own are just used to help visualise some of the information we want to show, but need to be substantiated with other information to drive home a clearer message.

